Predictions for detecting BHNS and other double compact objects binaries with LISA

### 2021-05-17 paper/tex/sections/discussion.tex
move distinguishing section to discussion
updates to methods after selma feedback

\subsection{Caveats}
\begin{itemize}
    \item Standard pop synth limitations
    \item Systemic kicks
    \item No models for high $Z$ values in Frankel
    \item Only a single disc for the Milky Way
\end{itemize}

\subsection{Can we distinguish them?}

>>>>

\subsection{Caveats}
Separate into three sections

GCs don't produce DNSs well but the other ones yes

Contribution of metal poor stars is not well known

\begin{itemize}
    \item Standard pop synth limitations
    \item Systemic kicks
    \item No models for high $Z$ values in Frankel
    \item Other channels
    \item Only a single disc for the Milky Way
\end{itemize}

\subsection{Identification of GW sources}
It is important to note that, though we present predictions for the detection rates of specific DCO types, the nature of the source may not be immediately apparent from the gravitational wave signal.

\subsubsection{Distinguishing from WDWD population}
\subsubsection{Separating BHBHs, BHNSs and NSNSs}

### 2021-05-17 paper/tex/sections/methods.tex
updates to methods after selma feedback

In order to estimate a detection rate of DCOs with statistical uncertainties, we create a series of random instances of the Milky Way, each populated with a subsample drawn (with replacement) from the synthesised binaries described in Section~\ref{sec:COMPAS_explained}.

\subsubsection{Milky Way model}\label{sec:mw_model}
We assume that the Milky Way is composed of a thin disc, thick disc and a central bar/bulge. For each of these components we use a separate lookback time, radial and vertical distribution, which we combine into a single model, weighting each component by its stellar mass. \citet{Licquia+2015} gives that the stellar mass of the bulge is $0.91 \times 10^{10} \unit{M_{\odot}}$ and the stellar mass of the disc is $5.17 \times 10^{10} \unit{M_\odot}$, which we split equally between the thin and thick disc. \todo{need citation for equal mass discs}

\textit{Star formation history:} For the thin and thick discs we use the star formation history given by \citet[][Eq.~2]{Frankel+2018},
\begin{equation}\label{eq:thin_disc_tau}
    p(\tau) \propto \exp \qty(-\frac{(\tau_m - \tau)}{\tau_{\rm SFR}}),
\end{equation}
where $\tau$ is the lookback time (the amount of time elapsed between the binary's zero-age main sequence and today), $\tau_m = 12 \unit{Gyr}$ is the assumed age of the Milky Way and $\tau_{\rm SFR} = 6.8 \unit{Gyr}$ is the star formation timescale. Following \citet{Frankel+2018}, we assume that there is a transition from thick disc star formation to thin disc star formation $8 \unit{Gyr}$ ago and apply a sharp transition at this point. This has been shown to still reproduce the Milky Way structure well despite its simplicity \citep{Sharma+2020} \todo{I could write this bit better}. Therefore, for the thin disc we normalise this distribution on the range $[0, 8) \unit{Gyr}$ and for the thick disc we normalise on the range $[8, 12] \unit{Gyr}$.

For the bulge, we base our distribution on the one presented in \citet[][Fig.~7]{Bovy+2019}. We adapt this distribution to account for the possible bias of the sample, where fainter older stars are not observed \todo{I don't understand this well enough to write}. We therefore approximate the distribution using a $\beta(2,3)$ distribution, shifted and scaled such that stars are only formed in the range $[6, 12] \unit{Gyr}$.

@@

\textit{Vertical distribution}: Similar to the radial distribution, we use the same single exponential distribution (but with different scale heights) for each component
\begin{equation}\label{eq:galaxy_z}
    p(\abs{z}) = \frac{1}{z_d} \exp\qty(-\frac{z}{z_d}),
\end{equation}
where $z$ is the height above the Galactic plane and $z_d$ is the scale height. We set $z_d = 0.3 \unit{kpc}$ for the thin disc \citep{McMillan+2011} and $z_d = 0.95 \unit{kpc}$ for the thick disc \citep{Bovy+2016}. For the bulge, we set $z_d = 1.5 \unit{kpc}$ such that the probability of forming a star approaches zero at approximately the co-rotation radius of the Galactic bar/bulge \citep{Bovy+2019}.

\textit{Metallicity-radius-time relation:} Most previous studies that predict a detection rate for LISA place binaries in the Milky Way independently of their age or evolution. We improve upon this as the first study to use an analytical model of the Milky Way that takes into account the galaxy's enrichment history. For this purpose, we apply the metallicity-radius-time relation from \citet{Frankel+2018}. The authors developed this model in order to measure the global efficiency of radial migration in the Milky Way and calibrated it using a sample of red clump stars measured with APOGEE \citep{Majewski+2017} and matched to stars in Gaia \citep{GaiaCollaboration+2016}.

The metallicity-radius-time relation is given by \citep[][Eq. 7]{Frankel+2018}
\begin{equation}\label{eq:galaxy_FeH}
    \begin{split}
        [{\rm Fe} / {\rm H}] (R, \tau) &= F_m + \nabla [{\rm Fe} / {\rm H}] R \\
        &- \qty(F_m + \nabla [{\rm Fe} / {\rm H}] R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} ) f(\tau),
    \end{split}
\end{equation}
where
\begin{equation}
    f(\tau) = \qty(1 - \frac{\tau}{\tau_m})^{\gamma_{[{\rm Fe} / {\rm H}]}},
\end{equation}
$F_m = -1 \unit{dex}$ is the metallicity of the gas at the center of the disk at $\tau = \tau_m$, $\nabla [{\rm Fe} / {\rm H}] = -0.075 \unit{kpc^{-1}}$ is the metallicity gradient, $R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} = 8.7 \unit{kpc}$ is the radius at which the present day metallicity is solar and $\gamma_{[{\rm Fe} / {\rm H}]} = 0.3$ set the time dependence of the chemical enrichment. We can convert this to the representation of metallicity that we use in this paper by applying \citep[e.g][]{Bertelli+1994}
\begin{equation}\label{eq:galaxy_FeH_to_Z}
    \log_{10} (Z) = 0.977 [{\rm Fe} / {\rm H}] + \log_{10}(Z_\odot).
\end{equation}

@@

\subsubsection{Combining population and galaxy synthesis}

>>>>

In order to estimate a detection rate of DCOs with statistical uncertainties, we create a series of random instances of the Milky Way, each populated with a subsample drawn (with replacement) from the synthesised binaries described in Section~\ref{sec:COMPAS_explained}. Most previous studies that predict a detection rate for LISA place binaries in the Milky Way independently of their age or evolution. We improve upon this as the first study to use an empirically-informed analytical model of the Milky Way that takes into account the galaxy's enrichment history by applying the metallicity-radius-time relation from \citet{Frankel+2018}. The authors developed this model in order to measure the global efficiency of radial migration in the Milky Way and calibrated it using a sample of red clump stars measured with APOGEE \citep{Majewski+2017} and matched to stars in Gaia \citep{GaiaCollaboration+2016}.

In Section~\ref{sec:mw_model}, we outline our model for the Milky Way with each of the various distributions and in Section~\ref{sec:combining_pop_gal} we explain how we combine our population of synthesised DCOs with the Milky Way model.

\subsubsection{Milky Way model}\label{sec:mw_model}

Our model for the Milky Way accounts for the thin disc, thick disc and central bar/bulge. For each of these components we use a separate star formation history, radial and vertical distribution, which we combine into a single model, weighting each component by its stellar mass. \citet{Licquia+2015} gives that the stellar mass of the bulge is $0.9 \times 10^{10} \unit{M_{\odot}}$ and the stellar mass of the disc is $5.2 \times 10^{10} \unit{M_\odot}$, which we split equally between the thin and thick disc. \todo{@Neige, do you have a suggestion for a citation for having equal mass thin/thick discs}.

\textit{Star formation history:} For the thin and thick discs we use the star formation history given by \citet[][Eq.~2]{Frankel+2018},
\begin{equation}\label{eq:thin_disc_tau}
    p(\tau) \propto \exp \qty(-\frac{(\tau_m - \tau)}{\tau_{\rm SFR}}),
\end{equation}
where $\tau$ is the lookback time (the amount of time elapsed between the binary's zero-age main sequence and today), $\tau_m = 12 \unit{Gyr}$ is the assumed age of the Milky Way and $\tau_{\rm SFR} = 6.8 \unit{Gyr}$ is the star formation timescale. Following \citet{Frankel+2018}, we assume that there is a transition from thick disc star formation to thin disc star formation $8 \unit{Gyr}$ ago and apply a sharp transition at this point. Despite its simplicity, this has been shown to reproduce the Milky Way age and abundance distributions well \citep{Sharma+2020} \todo{@Neige, is this justification correct/relaying the results of Sharma+2020 properly?}. Therefore, for the thin disc we normalise this distribution on the range $[0, 8] \unit{Gyr}$ and for the thick disc we normalise on the range $[8, 12] \unit{Gyr}$.

For the bulge, we base our distribution on the one presented in \citet[][Fig.~7]{Bovy+2019}. This model is fit using the data from APOGEE and Gaia as \citet{Frankel+2018}. \todo{@Neige, I need a sentence here to explain what biases we account for and so why we shift it. I recall it's something to do with fainter stars being missed and the bright ones not being part of the bulge but I can't articulate it very well, could you please help me with this?}. We therefore approximate the distribution using a $\beta(2,3)$ distribution, shifted and scaled such that stars are only formed in the range $[6, 12] \unit{Gyr}$.

@@

\textit{Vertical distribution}: Similar to the radial distribution, we use the same single exponential distribution (but with different scale heights) for each component
\begin{equation}\label{eq:galaxy_z}
    p(\abs{z}) = \frac{1}{z_d} \exp\qty(-\frac{z}{z_d}),
\end{equation}
where $z$ is the height above the Galactic plane and $z_d$ is the scale height. We set $z_d = 0.3 \unit{kpc}$ for the thin disc \citep{McMillan+2011} and $z_d = 0.95 \unit{kpc}$ for the thick disc \citep{Bovy+2016}. For the bulge, we set $z_d = 1.5 \unit{kpc}$ such that the probability of forming a star approaches zero at approximately the co-rotation radius of the Galactic bar/bulge \citep{Bovy+2019}. \todo{@Neige, did I justify the bulge/bar height correctly?}

\textit{Metallicity-radius-time relation:} The relation is given by \citep[][Eq. 7]{Frankel+2018}
\begin{equation}\label{eq:galaxy_FeH}
    \begin{split}
        [{\rm Fe} / {\rm H}] (R, \tau) &= F_m + \nabla [{\rm Fe} / {\rm H}] R \\
        &- \qty(F_m + \nabla [{\rm Fe} / {\rm H}] R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} ) f(\tau),
    \end{split}
\end{equation}
where
\begin{equation}
    f(\tau) = \qty(1 - \frac{\tau}{\tau_m})^{\gamma_{[{\rm Fe} / {\rm H}]}},
\end{equation}
$F_m = -1 \unit{dex}$ is the metallicity of the gas at the center of the disc at $\tau = \tau_m$, $\nabla [{\rm Fe} / {\rm H}] = -0.075 \unit{kpc^{-1}}$ is the metallicity gradient, $R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} = 8.7 \unit{kpc}$ is the radius at which the present day metallicity is solar and $\gamma_{[{\rm Fe} / {\rm H}]} = 0.3$ set the time dependence of the chemical enrichment. We can convert this to the representation of metallicity that we use in this paper by applying \citep[e.g][]{Bertelli+1994}
\begin{equation}\label{eq:galaxy_FeH_to_Z}
    \log_{10} (Z) = 0.977 [{\rm Fe} / {\rm H}] + \log_{10}(Z_\odot).
\end{equation}

Although \citet{Frankel+2018} only fit this model for the thin disc, we use this same metallicity-radius-time relation for each component of the galaxy. \citet{Sharma+2020} showed that using a simple continuous model for both the thin and thick disc the Milky Way abundance distributions well and hence we can apply this same relation to the thick disc. \citet{Bovy+2019} show in Fig.~7 that the abundance distribution and chemical tracks of the stars in the bulge/bar follow the same track as those outside and hence we can use the same metallicity-radius-time relation.

@@

\subsubsection{Combining population and galaxy synthesis}\label{sec:combining_pop_gal}
